---
title: "Fst SLiM"
output: html_document
---

```{r}
suppressMessages(library(tidyverse))
```


Function to calculate Fst

```{r}
calcFst <- function(p1, p2, N1, N2) {
q1 <- 1 - p1
q2 <- 1 - p2
Ntotal <- N1 + N2
p_bar <- ((p1*N1) + (p2*N2)) / (Ntotal)
q_bar <- ((q1*N1) + (q2*N2)) / (Ntotal)
Hexp1 <- 2 * p1 * q1
Hexp2 <- 2 * p2 * q2
Hs <- ((Hexp1*N1) + (Hexp2*N2))/(Ntotal)
Ht <- 2 * p_bar * q_bar
Fst <- (Ht-Hs) / Ht
return(Fst)
}
```


I limited the source population sizes to 2 diploid individuals because calling local ancestry on whole genome SLiM files can be very slow with too many roots.

```{r}
#vector of possible allele frequencies with N=2
p1_vector <- c(0, 0.25, 0.5, 0.75, 1)
p2_vector <- c(0, 0.25, 0.5, 0.75, 1)

#grid of possible pairs of allele frequencies for source populations
gs <- cross_df(list(p1=p1_vector, p2=p2_vector))
head(gs)
```

Calculate Fst for all pairs of source population allele frequencies

```{r}
Fst_vector <- mapply(FUN=calcFst,p1=gs$p1, p2=gs$p2, N1=4, N2=4)
gs$Fst <- Fst_vector

#all possible Fst values
levels(as.factor(gs$Fst))
```

I picked Fst values of 0, 0.25, 0.60, and 1 because they were simplest. I selected possible pairs of allele frequencies that gave each Fst value

```{r}
gs[which(gs$Fst==0),]
gs[which(gs$Fst==0.25),]
gs[which(gs$Fst==0.6),]
gs[which(gs$Fst==1),]
```

I picked the allele frequencies that would make the most sense to do a local ancestry outlier analysis for P1 ancestry. 

In the end, I ran 10 simulations each of:

Fst = 0; p1 = 0.5; p2 = 0.5
Fst = 0.25; p1 = 0.75; p2 = 0.25
Fst = 0.6; p1 = 1; p2 = 0.25
Fst = 1; p1 = 1; p2 = 0

```{r}
FST_results <- read.table(file = "/Users/iman/Dropbox (Duke Bio_Ea)/GoldbergLab/Datasets/PhilTrans_Fst_whole-autosome_ancestryprop.txt", header=TRUE)

Fst_results_grouped <- FST_results %>% group_by(as.factor(Fst)) %>% summarise(median=median(bene_rank))

Fst_results_grouped

names(Fst_results_grouped) <- c("Fst", "median_percentile_rank")

ggplot(ungroup(Fst_results_grouped), aes(x=as.numeric(as.character(Fst)), y=median_percentile_rank)) +
  geom_point(size=2) +
  geom_line(group=1) +
  labs(y="median percentile rank", x="Fst") +
  theme_classic() +
  theme(aspect.ratio = 1)
```

